---
title: "Prepare MODIS Data"
output: html_notebook
---

This code imports, cleans, and merges MODIS product data exported using the [AppEEARS](https://lpdaacsvc.cr.usgs.gov/appeears) web application.

Exported MODIS data for the period 2000-2016:

* **MOD11A2:** LST Day 1km & LST Night 1km (8 day)
* **MOD13Q1:** EVI 250m & NDVI 250m (16 day)
* **MOD15A2H:** fPAR 500m & LAI 500m (8 day)
* **MOD16A2:** LE 500m & PET 500m (8 day)
* **MOD17A2H:** GPP 500m (8 day)

Data was exported for 50 BAMS paper sites (.csv format, "ID", "Category", "LAT", "LONG")
```{r echo=FALSE, warning=FALSE, message=FALSE}
# clear workspace and get wd
rm(list=ls())
# change wd
setwd("C:/Users/Gavin McNicol/Box Sync/MODIS Data/AppEEARS Data/")
bams.sites <- read.csv("AppEEARS_sites.csv")
names(bams.sites)[c(3,4)] <- c("LAT", "LONG")
bams.sites
```


```{r echo=FALSE}
# libraries
# load tidyverse libraries
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(stringr)
library(lubridate)
library(RColorBrewer)
library(ggplot2)
library(data.table)
library(magrittr)
library(caret)
library(caTools)
library(ranger)
library(tibble)
```

9 MODIS products are downloaded in 5 .csv files:
```{r message=FALSE, warning=FALSE, echo=FALSE}
# change wd
setwd("C:/Users/Gavin McNicol/Box Sync/MODIS Data/AppEEARS Data/bams-sites")
# get site names from directory
site.names <- list.files(pattern = "\\.csv$",full.names=TRUE)
site.names
  # create list of products
modis <- list()
# create a function to read many csvs
read.many.csv <- function(i){
  read.csv(i, header=TRUE)
}
# execute function with site.names
modis <- lapply(site.names, read.many.csv)
modis
```
Simplify data by removing unnecssary fields:
```{r echo = FALSE}
lst.names <- names(modis[[1]][c(3,4,5,9,10,15,24)]) 
vi.names <- names(modis[[2]][c(3,4,5,9,10,14)])
lai.names <- names(modis[[3]][c(3,4,5,9,10,14)])
et.names <- names(modis[[4]][c(3,4,5,9,10,14)])
gpp.names <- names(modis[[5]][c(3,4,5,9,13)])
names.vec <- list(lst.names,vi.names, lai.names, et.names, gpp.names)
modis.simple <- list()
for (i in 1:length(modis)){
     modis.simple[[i]] <- as_tibble(modis[[i]][names.vec[[i]]])
     modis.simple[[i]]$Date <- as_date(modis.simple[[i]]$Date)
}
#split night and day LST; evi and NDVI; LAI and fPAR.
modis.simple[[6]] <- modis.simple[[1]] %>% select(Latitude, Longitude, Date, MOD11A2_006_LST_Night_1km, 
                                                  MOD11A2_006_QC_Night_MODLAND_Description)
modis.simple[[1]] <- modis.simple[[1]] %>% select(Latitude, Longitude, Date, MOD11A2_006_LST_Day_1km,
                                                  MOD11A2_006_QC_Day_MODLAND_Description)
names(modis.simple) <- c("lst_day","vi","lai","et","gpp","lst_night")
modis.simple
```
Look at what the data quality levels are, convert good to 1, keep only 1s
```{r echo= FALSE}
#convert levels to factor
levels(modis.simple[[1]][[5]]) <- as.factor(c(0,1,0))
levels(modis.simple[[2]][[6]]) <- as.factor(c(0,0,1,0))
levels(modis.simple[[3]][[6]]) <- as.factor(c(1,0))
levels(modis.simple[[4]][[6]]) <- as.factor(c(1,0))
levels(modis.simple[[5]][[5]]) <- as.factor(c(1,0))
levels(modis.simple[[6]][[5]]) <- as.factor(c(0,1,0))
#rename variables
names(modis.simple[[1]]) <- c("LAT", "LONG", "DATE", "LST_D", "QUALITY")
names(modis.simple[[2]]) <- c("LAT", "LONG", "DATE", "EVI", "NDVI", "QUALITY")
names(modis.simple[[3]]) <- c("LAT", "LONG", "DATE", "FPAR", "LAI", "QUALITY")
names(modis.simple[[4]]) <- c("LAT", "LONG", "DATE", "LE", "PET", "QUALITY")
names(modis.simple[[5]]) <- c("LAT", "LONG", "DATE", "GPP", "QUALITY")
names(modis.simple[[6]]) <- c("LAT", "LONG", "DATE", "LST_N", "QUALITY")
# modis.simple

modis.filtered <- list()
for (i in 1:length(modis.simple)){
  modis.filtered[[i]] <- as_tibble(modis.simple[[i]] %>% filter(QUALITY == 1))
}
modis.filtered
names(modis.filtered) <- c("lst_day","vi","lai","et","gpp","lst_night")
```
Make a plot of daytime LST to make sure things look OK
```{r}
ggplot(modis.filtered[[1]], aes(DATE, LST_D, col = as.factor(LAT))) +
  geom_point()
```
Create a dataframe for each product consisting of LAT, LONG, DATE and the product field (e.g. LST_D)
```{r}
lst_d.filtered <- modis.filtered$lst_day %>% mutate(LST_D = LST_D) %>% select(LAT, LONG, DATE, LST_D)
lst_n.filtered <- modis.filtered$lst_night %>% mutate(LST_N = LST_N) %>% select(LAT, LONG, DATE, LST_N)
evi.filtered <- modis.filtered$vi %>% mutate(EVI = EVI) %>% select(LAT, LONG, DATE, EVI)
ndvi.filtered <- modis.filtered$vi  %>% mutate(NDVI = NDVI) %>% select(LAT, LONG, DATE, NDVI)
fpar.filtered <- modis.filtered$lai  %>% mutate(FPAR = FPAR) %>% select(LAT, LONG, DATE, FPAR)
lai.filtered <- modis.filtered$lai  %>% mutate(LAI = LAI) %>% select(LAT, LONG, DATE, LAI)
gpp.filtered <- modis.filtered$gpp %>% mutate(GPP = GPP) %>% select(LAT, LONG, DATE, GPP) 
le.filtered <- modis.filtered$et %>% mutate(LE = LE) %>%  select(LAT, LONG, DATE, LE) 
pet.filtered <- modis.filtered$et %>% mutate(PET = PET) %>% select(LAT, LONG, DATE, PET) 
```
Create a full dataframe with each site name/latlong repeated n = length(2000-01-01 to 2018-12-31) (MODIS data window).
Then right_join each modis product to the site name/all dates tibble to produce a complete dataframe of modis data.
```{r}
full.date <- as_tibble(rep(as_date(c(as_date("2000-01-01"):as_date("2018-12-31"))),50))
names(full.date) <- "DATE"
full.date <- full.date %>% add_column(ID = rep(bams.sites$ID, each = length(full.date$DATE)/50)) %>% 
                              add_column(LAT = rep(bams.sites$LAT, each = length(full.date$DATE)/50)) %>%
                                add_column(LONG = rep(bams.sites$LONG, each = length(full.date$DATE)/50))


modis.full <- right_join(full.date, lst_d.filtered, by =c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, lst_n.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, evi.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, ndvi.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, fpar.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, lai.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, gpp.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, le.filtered, by=c("DATE","LAT","LONG"))
modis.full <- right_join(modis.full, pet.filtered, by=c("DATE","LAT","LONG"))
modis.full <- left_join(modis.full, bams.sites, by=c("LAT", "LONG"))
modis.full <- modis.full %>% select(ID = ID.y, Category, -ID.x, LAT, LONG, DATE, LST_D, LST_N, EVI, NDVI, FPAR, LAI, GPP, LE, PET)

str(modis.full)
modis.full
```

Finished! Save as .csv
```{r warning=FALSE, message=FALSE}
# set wd
setwd("C:/Users/Gavin McNicol/Box Sync/MODIS Data/AppEEARS Data")
write.csv(modis.full, "bams_sites_MODIS_full.csv")
```




